---
title: "Stat 333 Project Results"
author: "Jonathan Morris"
date: "2024-05-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages:
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(interactions)
library(usmap)
```

# Load the data:
```{r}
data <- read.csv("https://raw.githubusercontent.com/ajcarney/STAT333Project/master/clean_data/data.csv")
data <- data %>% select(-Perc_Dem, -Perc_Rep) %>% drop_na()
```


# Run original model:
```{r}
fit1 <- lm(ev_market_penetration ~ Income_adjusted*gas_regular + gas_tax + n_charging_locations + log_population + avg_temp + Perc_Adults_Bachelor_Or_Above, data = data)
summary(fit1)
```

# Plot original model (with jitter):
```{r}
ggplot(data, aes(x=Income_adjusted, y = ev_market_penetration)) +
  geom_point(aes(color = gas_regular), position = position_jitter(width = 0.05, height = 0.05)) +
  geom_smooth(method = "lm") +
  ggtitle("Median Household Income Associated With EV Market Penetration") +
  labs(x = "Median Household Income / RPP ($)", y = "Electric Vehicle Market Penetration") + 
  scale_color_continuous(name = "Average Gas Price ($/g)") + 
  scale_color_gradient(name = "Average Gas Price ($/g)",
                       low = "pink",  # lighter color
                       high = "darkred")
```

# Show diagnostic plots:
```{r}
plot(fit1)
```

# Model with gas price non-imputated (Aiden can fill in):

# Run model with log income:
```{r}
fit2 <- lm(ev_market_penetration ~ log_income*gas_regular + gas_tax + n_charging_locations + log_population + avg_temp + Perc_Adults_Bachelor_Or_Above, data = data)
summary(fit2)
```
#Diagnostic plots for log_income:
```{r}
plot(fit2)
```


#Plot log_income
```{r}
ggplot(data, aes(x=log_income, y = ev_market_penetration)) +
  geom_point(aes(color = gas_regular),position = position_jitter(width = 0.05, height = 0.05)) +
  geom_smooth(method = "lm") +
  ggtitle("Median Household Income Associated With EV Market Penetration") +
  labs(x = "Logged Median Household Income ($)", y = "Electric Vehicle Market Penetration") + 
  scale_color_continuous(name = "Average Gas Price ($/g)") + 
  scale_color_gradient(name = "Average Gas Price ($/g)",
                       low = "pink",  # lighter color
                       high = "darkred")
```

# Run the model with log income adjusted:
```{r}
data <- data %>% drop_na();
fit3 <- lm(ev_market_penetration ~ log_income_adjusted*gas_regular + gas_tax + n_charging_locations + log_population + avg_temp + Perc_Adults_Bachelor_Or_Above, data = data)
summary(fit3)
```
```{r}


data2 <- data %>% mutate(
        resids = resid(fit3)
    )

plot_usmap(data = data2, values="resids", color = "grey", size = .25, linewidth=0.1) +
  scale_fill_gradient(low = "pink", high = "darkred", na.value = "transparent") +
  theme(legend.position = "right")
#plot(fit3)
```

# Plot log_income / RPP
### Shows a negative slope because the values were logged as a ratio. think of these results as the same but just flipped.
```{r}
ggplot(data, aes(x=log_income_adjusted, y = ev_market_penetration)) +
  geom_point(aes(color = gas_regular), position = position_jitter(width = 0.01, height = 0.01)) +
  geom_smooth(method = "lm") +
  ggtitle("Median Household Income Associated With EV Market Penetration") +
  labs(x = "Log (Median Household Income / RPP)", y = "Electric Vehicle Market Penetration") + 
  scale_color_continuous(name = "Average Gas Price ($/g)") + 
  scale_color_gradient(name = "Average Gas Price ($/g)",
                       low = "pink",  # lighter color
                       high = "darkred")
```
